---
title: "PS0014-Figur-1"
output: github_document
Author: "Prachi Sao"
Date: "2021-06-30"
organization:"PAUKEN LAB, MD ANDERSON CANCER CENTER"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## load library
```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(dittoSeq)
set.seed(1) #set a seed for reproducibility
options(future.globals.maxSize = +Inf)
```

## load colors
```{r}  
<!-- # set colors for plots  -->
my_colors_cluster <- c('red','black','blue','green','orange','yellow','brown','magenta','pink','gray')
my_colors_sample <- c('purple', 'tan','lightblue','lightgreen','maroon','lightgreen','lightpink','yellow',"orange","lightblue")
my_colors_cp_status =c("blue", "gray90")
```

## Load Data
```{r}
no_tcr_final = readRDS("Results /RDS/PS00111-cd8_cca_integrated_res04_final_04_22_2024_final.rds")
names(no_tcr_final@meta.data)
```

## Figure 1B
```{r }

Fig.1.B1 = dittoDimPlot(no_tcr_final, "PDCD1", max.color = "#de2d26", min.color = "#fee0d2",
             opacity = 0.7,
              main = "PDCD1 Transcript",
             legend.size = 8)
Fig.1.B1 + NoAxes()
# Create the plot without borders
Fig.1.B2 = dittoDimPlot(no_tcr_final, "CTLA4", max.color = "#de2d26", min.color = "#fee0d2",
             opacity = 0.7,
              main = "CTLA4 Transcript",
             legend.size = 8)
Fig.1.B2+ NoAxes()
```

## Figure 1C
```{r }
 
 Idents(no_tcr_final) = "cp_status"
# Assuming you have a Seurat object named 'no_tcr_final'
# First, create a vector of colors based on the number of cells in each cluster
cell_counts <- table(no_tcr_final$cp_status)
gray_color <- "gray90"  # Color for more cells
blue_color <- "blue"    # Color for fewer cells

# Assign colors based on cell counts
cluster_colors <- ifelse(cell_counts > median(cell_counts), gray_color, blue_color)

# Now create the DimPlot with custom colors
Fig.1.C = DimPlot(no_tcr_final, label = F, cols = cluster_colors)
Fig.1.C
```

## Figure 1D
```{r }
# no_tcr_final = readRDS("Results /RDS/PS00111-cd8_cca_integrated_res04_final_04_22_2024_final.rds")
# Idents(no_tcr_final) <- "seurat_cluster.04"
# levels(Idents(no_tcr_final))

# no_tcr_final <- RenameIdents(no_tcr_final,
#                     "0" = "terminally exhausted, cytotoxic-like, TRM-like TEX",
#                     "1" = "Granzyme K+ effector-memory-like",
#                     "2" = "activated, effector-like glycolytic TEX",
#                     "3" = "Naive and/or TCM-like",
#                     "4" = "CXCL13+ TCF7+ glycolytic TPEX-like",
#                     "5" = "mixed T cells/non-T cells",
#                     "6" = "effector-like",
#                     "7" = "cycling TEX",
#                     "8" = "IFN stim'd TEX",
#                     "9" = "non-T cells")

Fig.1.D = DimPlot(no_tcr_final, label = F, cols = my_colors_cluster, label.color = "black", repel = T, na.value = "grey50", label.size = 2, seed = 1) + 
                        theme(legend.position = "bottom", legend.box = "horizontal") + 
                        guides(col = guide_legend(nrow = 2))

# Print the customized plot
print(Fig.1.D)
```

## Figure 1E
```{r }
 
Fig.1.E = dittoBarPlot(no_tcr_final, var = "seurat_cluster.04", group.by = "cp_status",  color.panel = my_colors_cluster, legend.show = F, xlab = NULL, ylab ="Percentage of cell per seurat cluster") + coord_flip()
Fig.1.E

```

###################### gene signatures ############################
```{r}
#This code is to look at the genesignatures from litrature and halmar database 
#"Several gene signatures are read from an Excel file named “TableS4_SupplementGeneSigs.xlsx” located Raw_data folder 
# These gene signatures include: SadeFeldman_CD8_B, SadeFeldman_CD8_G, Trm_signature, cell_cycle_all_strong, Glycolytic
 ```


##load Excel file
```{r}
SadeFeldman_CD8_B <- read_excel("Raw_data/TableS4_SupplementGeneSigs.xlsx")$'Sade Feldman CD8 B (source Sade-Feldman et al. 2018)'
length(SadeFeldman_CD8_B)
SadeFeldman_CD8_G <- read_excel("Raw_data/TableS4_SupplementGeneSigs.xlsx", n_max = 34)$'Sade Feldman CD8 G (source Sade-Feldman et al. 2018)'
length(SadeFeldman_CD8_G)
Trm_signature<- read_excel("Raw_data/TableS4_SupplementGeneSigs.xlsx", n_max = 143)$'TRM (source Han et al. 2021)'
cell_cycle_all_strong <- read_excel("Raw_data/TableS4_SupplementGeneSigs.xlsx", n_max = 96)$'Cell Cycle (source Kowalczyk et al. 2015)'
Glycolytic <- read_excel("Raw_data/TableS4_SupplementGeneSigs.xlsx", n_max = 200)$'Glycolytic (source HALLMARKS database from MSigDB)'
```
## add gene signatures to the Seurat object using add module score
```{r}
# SadeFeldman_CD8_B
# no_tcr_final <- AddModuleScore(no_tcr_final,features = list(SadeFeldman_CD8_B),name="SadeFeldman_CD8_B")
# no_tcr_final <- AddModuleScore(no_tcr_final,features = list(SadeFeldman_CD8_G),name="SadeFeldman_CD8_G")
# no_tcr_final <- AddModuleScore(no_tcr_final,features = list(Trm_signature), name="Trm_signature")
# no_tcr_final <- AddModuleScore(no_tcr_final,features = list(cell_cycle_all_strong),name="cell_cycle_all_strong")
# no_tcr_final <- AddModuleScore(no_tcr_final,features = list(Glycolytic),name="Glycolytic")
```

## Figure 1FGHI
```{r}
SadeFeldman_CD8_B_plot =FeaturePlot(no_tcr_final,features = "SadeFeldman_CD8_B1", label = F, repel = TRUE) +
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu"))) 
SadeFeldman_CD8_B_plot            
SadeFeldman_CD8_G_plot =FeaturePlot(no_tcr_final,features = "SadeFeldman_CD8_G1", label = F, repel = TRUE) +
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))  
SadeFeldman_CD8_G_plot 
Trm_signature_plot = FeaturePlot(no_tcr_final,features = "Trm_signature1", label = F, repel = TRUE) +
                             scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))           
Trm_signature_plot
cell_cycle_all_strong_plot = FeaturePlot(no_tcr_final,features = "cell_cycle_all_strong1", label = F, repel = TRUE) +
                             scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))     
cell_cycle_all_strong_plot


Figure.1FGHI = grid.arrange(SadeFeldman_CD8_B_plot, SadeFeldman_CD8_G_plot, Trm_signature_plot, cell_cycle_all_strong_plot , 
ncol =2)
Figure.1FGHI
```

```{r}
SadeFeldman_CD8_B_plotv3 = VlnPlot_scCustom(seurat_object = no_tcr_final, features = "SadeFeldman_CD8_B1",group.by = "cp_status", plot_boxplot = TRUE)
SadeFeldman_CD8_B_plotv3+ stat_compare_means()
SadeFeldman_CD8_Gv3 = VlnPlot_scCustom(seurat_object = no_tcr_final, features = "SadeFeldman_CD8_G1",group.by = "cp_status", plot_boxplot = TRUE)
SadeFeldman_CD8_Gv3+ stat_compare_means()
Trm_signaturev3 = VlnPlot_scCustom(seurat_object = no_tcr_final, features = "Trm_signature1",group.by = "cp_status", plot_boxplot = TRUE)
Trm_signaturev3+ stat_compare_means()
cell_cycle_all_strongv3 = VlnPlot_scCustom(seurat_object = no_tcr_final, features = "cell_cycle_all_strong1",group.by = "cp_status", plot_boxplot = TRUE)
cell_cycle_all_strongv3+ stat_compare_means()


pval.SadeFeldman_CD8_B <- wilcox.test(SadeFeldman_CD8_B1 ~ ident,  data = as.data.frame(SadeFeldman_CD8_B_plotv3$data))$p.value
    pval.SadeFeldman_CD8_B
pval.SadeFeldman_CD8_g <- wilcox.test(SadeFeldman_CD8_G1 ~ ident,  data = as.data.frame(SadeFeldman_CD8_Gv3$data))$p.value
    pval.SadeFeldman_CD8_g
pval.trm = wilcox.test(Trm_signature1 ~ ident,  data = as.data.frame(Trm_signaturev3$data))$p.value
    pval.trm
pval.cc = wilcox.test(cell_cycle_all_strong1 ~ ident,  data = as.data.frame(cell_cycle_all_strongv3$data))$p.value
    pval.cc
pval.gly = wilcox.test(Glycolytic1 ~ ident,  data = as.data.frame(Glycolyticv3$data))$p.value
    pval.gly

grid.arrange(SadeFeldman_CD8_B_plotv3, SadeFeldman_CD8_Gv3, Trm_signaturev3, cell_cycle_all_strongv3 ,
ncol =2)
```