---
title: "PS0012-Figur-S1"
output: github_document
Author: "Prachi Sao"
Date: "2021-06-30"
organization:"PAUKEN LAB, MD ANDERSON CANCER CENTER"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This R Markdown document is used for visualizing single-cell RNA-seq data using various R packages. 
Figure S1:  
(B) UMAPs showing the distribution of the samples derived from the 7 patients included in the single cell RNA seq dataset. 
(C) Quantification of the absolute number of cells per sample contributing to each of the 10 transcriptional clusters identified. 
(D) Quantification of the frequency of each sample per cluster for the 10 transcriptional clusters identified. 
(E) UMI distribution per sample per cluster for the single cell RNA seq dataset. 

Setup: Loads necessary libraries for data manipulation, visualization, and analysis.
Load Colors: Defines custom color palettes for plots.
Figure S1.B: Generates a UMAP plot colored by clusters and split by sample.
Figure S1.C: Creates a stacked bar plot showing cluster sizes by sample.
Figure S1.D: Produces a proportional bar plot showing the distribution of samples within each cluster.
Figure S1.E: Displays a bar plot of UMI counts per cluster per sample, with side-by-side bars for each cluster.

## Including Code

You can include R code in the document as follows:

```{r}
library(knitr)
library(Seurat, quietly = TRUE) # Single-cell RNA-seq analysis
library(Azimuth, quietly = TRUE) # Reference-based single-cell RNA-seq analysis
library(presto, quietly = TRUE) # Fast differential expression analysis
library(dplyr, quietly = TRUE) # Data manipulation
library(cowplot, quietly = TRUE) # Plotting
library(viridis, quietly = TRUE) # Color palettes
library(gridExtra, quietly = TRUE) # Miscellaneous functions for "grid" graphics
library(data.table, quietly = TRUE) # Data manipulation
library(tibble, quietly = TRUE) # Data frames
library(Matrix, quietly = TRUE) # Sparse and dense matrix classes and methods
library(ggplot2, quietly = TRUE) # Data visualization
library(patchwork, quietly = TRUE) # Combine plots
library(grid, quietly = TRUE) # The Grid Graphics Package
library(RColorBrewer, quietly = TRUE) # Color palettes
library(clustree, quietly = TRUE) # Cluster trees
library(pheatmap, quietly = TRUE) # Pretty heatmaps
library(msigdbr, quietly = TRUE) # MSigDB gene sets
library(purrr, quietly = TRUE) # Functional programming tools
library(clusterProfiler, quietly = TRUE) # Statistical analysis and visualization of functional profiles
library(scater, quietly = TRUE) # Single-cell RNA-seq analysis
library(MatrixGenerics, quietly = TRUE) # S4 generic functions for matrix objects
library(tidyverse, quietly = TRUE) # Data science packages
library(ComplexHeatmap, quietly = TRUE) # Complex heatmaps
library(circlize, quietly = TRUE) # Circular visualization
library(scCustomize, quietly = TRUE) # Customization for single-cell analysis
library(qs, quietly = TRUE) # Quick serialization of R objects
library(dittoSeq, quietly = TRUE) # Visualization of single-cell and bulk RNA-seq data
library(ggpubr, quietly = TRUE) # 'ggplot2' based publication ready plots
library(reshape2, quietly = TRUE) # Flexibly reshape data
require(fgsea, quietly = TRUE) # GSEA analysis
require(useful, quietly = TRUE) # Corner function to peek at data.frames
require(readxl, quietly = TRUE) # Reading Excel files
require(msigdbr, quietly = TRUE) # MSigDB gene sets

set.seed(1) #set a seed for reproducibility
```

## load colors
```{r}  
<!-- # set colors for plots  -->
my_colors_cluster <- c('red','black','blue','green','orange','yellow','brown','magenta','pink','gray')
my_colors_sample <- c('purple', 'tan','lightblue','lightgreen','maroon','lightgreen','lightpink','yellow',"orange","lightblue")
my_colors_cp_status =c("blue", "gray90")
```

## Figure S1 A
# The bulk RNA-seq data were re-analyzed to derive gene signatures for CPHi and CPLo TILs. 
# Samples were flash frozen and sequenced to a 25M read depth using an Illumina platform. 
# Salmon v1.2.1 (RRID:SCR_017036) quantified transcripts from FASTQ files using the Ensembl GRCh38 reference genome (RRID:SCR_002344). 
# The counts were analyzed with DESeq2 v1.32 (RRID:SCR_000154), using the DESeqDataSetFromTximport command. 
# Gene set enrichment analysis was performed using GSEA 4.2.3 (RRID:SCR_003199).
# Top 200 up-regulated genes were identified and used for this analysis.

##Figure S1.B
```{r }
# no_tcr_final = readRDS("Results /RDS/PS00111-cd8_cca_integrated_res04_final_04_22_2024_final.rds")
#  p <- DimPlot(no_tcr_final, group.by = "sample") + ggtitle("color by samples")
#   print(p)
Figure S1.B <- DimPlot(no_tcr_final, group.by = "seurat_cluster.04", split.by = "sample", ncol = 3) + 
    scale_color_manual(values = my_colors_cluster) + 
    ggtitle("color by clusters")
print(Figure S1.B)
```

##Figure S1.C
```{r }
Figure S1.C <- no_tcr_final@meta.data %>%
    group_by(sample, seurat_cluster.04) %>%
    tally() %>%
    ggplot(aes(x = seurat_cluster.04, y = n, fill = as.factor(sample))) +
    geom_bar(stat = "identity") +  # Stacked bar plot
    theme_classic() +  # Use minimal theme
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = my_colors_sample) +
    labs(title = "Cluster Sizes by Sample")
print(Figure S1.C)
```

##Figure S1.D

```{r }
Idents(no_tcr_final) = "sample"
# Assuming 'no_tcr_final' is your Seurat object
pt <- table(Idents(no_tcr_final), no_tcr_final$seurat_cluster.04)
pt <- as.data.frame(pt)
head(pt)
pt$Var1 <- as.character(pt$Var1)

# Create a custom color palette with ten shades of blue
my_colors_stackedbarplot <- colorRampPalette(brewer.pal(10, "Blues"))(10) 
Figure S1.D  = ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
    theme_bw(base_size = 9) +
    geom_col(position = "fill", width = 0.5, color = "black") +  # Add border to each bar
    xlab("Cluster Number") +
    ylab("Proportion of Each Sample within Each Cluster") +theme_classic() +
    scale_fill_manual(values = my_colors_stackedbarplot) +
    theme(legend.title = element_blank(), legend.position = "bottom") +
    ggtitle("Distribution: Sample per Cluster") +
    coord_flip() +
    guides(fill = guide_legend(nrow = 3, byrow = TRUE))
print(Figure S1.D)
```

##Figure S1.E
```{r }
p <- no_tcr_final@meta.data %>%
    group_by(sample, seurat_cluster.04) %>%
    tally() %>%
    ggplot(aes(x = sample, y = n, fill = as.factor(seurat_cluster.04))) +
    geom_bar(stat = "identity", position = "dodge") +  # Display clusters side by side
    geom_text(aes(label = n), position = position_dodge(width = .9), vjust = -0.5, size = 3.5) +  # Adjust text position and size
    theme_classic() +  # Use classic theme
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = my_colors_cluster) +
    labs(x = "Sample ID", y = "Number of UMI's") +  # Add x and y labels
    theme(text = element_text(size = 15)) +  # Set font size to 9
    ggtitle("UMI Distribution per Cluster per Sample") +
    theme(legend.position = c(0.05, 0.95), legend.justification = c("left", "top")) +  # Position legend inside plot at top left
    guides(fill = guide_legend(nrow = 2))  # Keep legend in 2 rows
print(p)
```
