---
title: "PS003-Figur-S5-validation-sadefeldman"
output: github_document
Author: "Prachi Sao"
Date: "2021-06-30"
organization:"PAUKEN LAB, MD ANDERSON CANCER CENTER"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data_S_tcell.cd8 = readRDS("./Results/RDS/PS003-Figur-S4-validation-sadefeldman-cd8_tcells_cd8.rds")
```

# Figure S4 A
```{r}
sum(data_S_tcell.cd8[["RNA"]]$data < 0)

co_expression = c("PDCD1", "HAVCR2", "TIGIT", "LAG3", "CTLA4", "CD160")
data_S_tcell.cd8[["RNA"]]$data[co_expression,]

new_metadata <- data.frame(coexpression = rowSums(as.data.frame(t(as.data.frame(data_S_tcell.cd8[["RNA"]]$data[co_expression,]))) > 0))
data_S_tcell.cd8<- AddMetaData(data_S_tcell.cd8, new_metadata)

#plot coexpression
my_colors_stackedbarplot <- colorRampPalette(brewer.pal(7, "Blues"))(7)

dittoDimPlot(data_S_tcell.cd8, "coexpression", split.by = "coexpression",  color.panel = colorRampPalette(brewer.pal(7, "Blues"))(7), 
  colors = seq_along(color.panel))

```

# Figure S4 B
```{r}
Idents(data_S_tcell.cd8) = "sub_cluster_cd8.3"

my_colors_stackedbarplot <- colorRampPalette(brewer.pal(7, "Blues"))(7) 
Idents(data_S_tcell.cd8) = "coexpression"

pt_co  = table(Idents(data_S_tcell.cd8), data_S_tcell.cd8$sub_cluster_cd8.3)
pt_co <- as.data.frame(pt_co)
pt_co$Var1 <- as.character(pt_co$Var1)
pt_co$Var1 

barplo_co = ggplot(pt_co, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Coexpression") +
  labs(fill='IRs') +
  ylab("Percentage of cell per seurat cluster") +
  #scale_fill_manual(values = brewer.pal(12, "Paired")) +
  scale_fill_manual(values = my_colors_stackedbarplot) +
  theme(legend.title = element_blank()) +theme_classic()
  barplo_co 

```

# Figure S4 C
```{r}
FigS1D_gene_list = c("PDCD1", "HAVCR2", "LAG3", "BTLA", "CD160", "CTLA4", "TIGIT", "ENTPD1", "CD244", "PROCR", "PDPN", "CD200",
"TNFRSF9", "TNFRSF18", "CD27", "CD28", "CD226", "CD40LG",
"TCF7", "CCR7", "SELL", "IL7R", "LEF1", "KLF2",
"GZMB", "PRF1", "GZMA", "GZMK", "IFNG", "TNF", "IL2",
"ITGAE", "ITGA1", "CDH1", "CXCR6",
"NR4A1", "EOMES", "HIF1A", "TOX", "ZEB2", "ID2", "ID3", "BHLHE40","CXCL13",
"CXCR3", "CX3CR1", "KLRG1", "CD101", "ENO1", "GAPDH", "ALDOA", "PGK1", "LDHA")

#TXB21 is not available in data

aggregate_FigS1D_gene_list_df<- as.data.frame(AggregateExpression(data_S_tcell.cd8, features = FigS1D_gene_list, group.by ="sub_cluster_cd8.3", return.seurat = F, assays = "RNA"))


#organize the dataframe rows acording to gene_list
aggregate_FigS1D_gene_list_df<- data.matrix(aggregate_FigS1D_gene_list_df)[match(FigS1D_gene_list, rownames(data.matrix(aggregate_FigS1D_gene_list_df))),]

colnames(aggregate_FigS1D_gene_list_df) = c(levels(data_S_tcell.cd8))
scaled.mat2 <- t(scale(t(aggregate_FigS1D_gene_list_df)))
FigS1D.mat = data.matrix(scaled.mat2)[match(FigS1D_gene_list, rownames(data.matrix(scaled.mat2))),]
data.matrix(FigS1D.mat)
library(pheatmap)

input.hm.filter.breaks <- t(scale(t(FigS1D.mat )))
quantile.range <- quantile(t(input.hm.filter.breaks), probs = seq(0, 1, 0.1))
paletteBreaks <- quantile.range
paletteLength <- length(paletteBreaks)-1
col.pal <- RColorBrewer::brewer.pal(paletteLength, "RdBu")
col.pal <- rev(col.pal)

pheatmap(
    mat               = FigS1D.mat,
    breaks            = paletteBreaks,
    cluster_cols      = FALSE,
    cluster_rows      = FALSE,
    color             = col.pal,
    border_color      = NA,
    show_colnames     = T,
    show_rownames     = T,
    drop_levels       = TRUE,
    fontsize          = 14,
    main              = "figS4C",
      # Add this line to reorder the heatmap based on column
)
```


## Figure S5 D & E
##Natural break is at 0.3, hence cells expressing PDCD1 and CTLA4 above 0.3 were selected as cp high
```{r}
```{r}
gene1 <- "PDCD1"
gene2 <- "CTLA4"


pdcData <- FetchData(data_S_tcell.cd8, vars = gene1)
head(pdcData)

# Plot histogram for PDCD1
pdcd1_th2 = ggplot(pdcData, aes(x = log2(1 + log(1 + pdcData[, 1])))) +
  geom_histogram(binwidth = 0.2, fill = "blue", color = "black") +
  labs(x = paste("Log2(1+Log(1+x)) of", gene1, "counts"), y = "Frequency") +
  scale_x_continuous(breaks = c(0,0.1, 0.2,0.3,0.4, 0.5, 1, 1.5)) +
  theme_classic()

pdcd1_th2 
 # Filter data where PDCD1 expression is greater than 0.5
filtered_pdcData <- pdcData[pdcData[, 1] > 0.5, ]
length(filtered_pdcData)

# Extract expression data for CTLA4
ctlaData <- FetchData(data_S_tcell.cd8, vars = gene2)
head(ctlaData)
# Plot histogram for CTLA4
ctla4_th = ggplot(ctlaData, aes(x = log2(1 + log(1 + ctlaData[, 1])))) +
  geom_histogram(binwidth = 0.2, fill = "blue", color = "black") +
  labs(x = paste("Log2(1+Log(1+x)) of", gene2, "counts"), y = "Frequency") +
  scale_x_continuous(breaks = c(0,0.1, 0.2,0.3,0.4, 0.5, 1, 1.5)) +
  theme_classic()
  
ctla4_th 
 # Filter data where CTLA4 expression is greater than 0.5
filtered_ctlaData <- ctlaData[ctlaData[, 1] > 0.5, ]
length(filtered_ctlaData)
```

## Figure S5 F
```{r}
my_colors_cluster <- c("#F8766D", "#B79F00", "#00BA38" ,"#00BFC4" ,"#619CFF", "#F564E3")

CPhi_bulkRNAseq_v4200 = VlnPlot_scCustom(seurat_object = data_S_tcell.cd8, features = "CPhi_bulkRNAseq1", plot_boxplot = T, colors_use = my_colors_cluster)+labs(y = "Genescore")+
ggtitle("CPhi Gene Signatures from Bluk-RNASeq")+
theme(axis.text.x = element_text(size = 6, angle = 65),  # Adjust x-axis label font size
        axis.text.y = element_text(size = 5, angle = 90))+  # Adjust y-axis label font size
theme(legend.position = "none") 
print(CPhi_bulkRNAseq_v4200)

CPlo_bulkRNAseq_v5200 = VlnPlot_scCustom(seurat_object = data_S_tcell.cd8, features = "CPlo_bulkRNAseq1",plot_boxplot = T, colors_use = my_colors_cluster)+labs(y = "Genescore")+
ggtitle("CPlo Gene Signatures from Bluk-RNASeq")+
theme(axis.text.x = element_text(size = 6, angle = 65),  # Adjust x-axis label font size
        axis.text.y = element_text(size = 5, angle = 90))+  # Adjust y-axis label font size
theme(legend.position = "none") 
print(CPlo_bulkRNAseq_v5200)

```