---
title: "PS0014-Figur-3"
output: github_document
Author: "Prachi Sao"
Date: "2021-06-30"
organization:"PAUKEN LAB, MD ANDERSON CANCER CENTER"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load colors
```{r}  
<!-- # set colors for plots  -->
my_colors_cluster <- c('red','black','blue','green','orange','yellow','brown','magenta','pink','gray')
my_colors_sample <- c('purple', 'tan','lightblue','lightgreen','maroon','lightgreen','lightpink','yellow',"orange","lightblue")
my_colors_cp_status =c("blue", "gray90")
```

## Load Data
```{r}
no_tcr_final = readRDS("Results /RDS/PS00111-cd8_cca_integrated_res04_final_04_22_2024_final.rds")
names(no_tcr_final@meta.data)
```

## Figure 3A & 3B
# Figure 3A and 3B are part of topic modelling and will be a separate plot named "PS001-Figure-S8-LDA-Topic-Models"

## Figure 3C
```{r}
Glycolytic <- read_excel("Raw_data/TableS4_SupplementGeneSigs.xlsx", n_max = 200)$'Glycolytic (source HALLMARKS database from MSigDB)'

Figure.3C.1 = FeaturePlot(no_tcr_final,features = "Glycolytic1", label = F, repel = TRUE) +
                             ggplot2::scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))     
Figure.3C.1

Figure.3C.2 = VlnPlot_scCustom(seurat_object = no_tcr_final, features = "Glycolytic1",group.by = "cp_status", plot_boxplot = TRUE)
Figure.3C.2#+ stat_compare_means()

pval.gly = wilcox.test(Glycolytic1 ~ ident,  data = as.data.frame(Glycolyticv3$data))$p.value
    pval.gly
```
## Figure 3D
## Figure 3D
# The bulk RNA-seq data were re-analyzed to derive gene signatures for CPHi and CPLo TILs. 
# Samples were flash frozen and sequenced to a 25M read depth using an Illumina platform. 
# Salmon v1.2.1 (RRID:SCR_017036) quantified transcripts from FASTQ files using the Ensembl GRCh38 reference genome (RRID:SCR_002344). 
# The counts were analyzed with DESeq2 v1.32 (RRID:SCR_000154), using the DESeqDataSetFromTximport command. 
# Gene set enrichment analysis was performed using GSEA 4.2.3 (RRID:SCR_003199).
# Top 200 up-regulated genes were identified and used for this analysis.


## Figure 3E
```{r}
e4 <- FetchData(no_tcr_final, vars = c("Glycolytic1","Trm_signature1","seurat_cluster.04") )
head(e4)

# Calculate the median values for each cluster
values <- aggregate(cbind(Glycolytic1, Trm_signature1) ~ seurat_cluster.04, data = e4, FUN = mean )
Figure.3E1 = ggplot(values, aes(x = Trm_signature1, y = Glycolytic1)) +
    geom_point(size = 5, shape = 16) +
    labs(y = "Glycolysis Signature Score", x = "TRM (Han et al. 2021) Signature Score",
             title = "Average Signature Scores Per Seurat Cluster") +
    theme_classic() +
    geom_text(aes(label = seurat_cluster.04), hjust = 0.5, vjust = -0.5, size = 8) +
    theme(axis.title.x = element_text(size = 14),
                axis.title.y = element_text(size = 14))
Figure.3E1
```

```{r}
e5 <- FetchData(no_tcr_final, vars = c("Glycolytic1","SadeFeldman_CD8_B1","seurat_cluster.04") )
head(e4)

# Calculate the median values for each cluster
values <- aggregate(cbind(Glycolytic1, SadeFeldman_CD8_B1) ~ seurat_cluster.04, data = e5, FUN = mean )
Figure.3E2 = ggplot(values, aes(x = SadeFeldman_CD8_B1, y = Glycolytic1)) +
  geom_point(size = 3, shape = 16) +
  labs(y = "Glycolysis (HALLMARKS) Signature Score", x = "SadeFeldman_CD8_B1 Signature Score",
       title = "Average Signature Scores Per Seurat Cluster") +
theme_classic() +
    geom_text(aes(label = seurat_cluster.04), hjust = 0.5, vjust = -0.5, size = 8) +
    theme(axis.title.x = element_text(size = 14),
                axis.title.y = element_text(size = 14))
Figure.3E2
```

```{r}
e6 <- FetchData(no_tcr_final, vars = c("Glycolytic1","cell_cycle_all_strong1","seurat_cluster.04") )
head(e4)

# Calculate the median values for each cluster
values <- aggregate(cbind(Glycolytic1,cell_cycle_all_strong1) ~ seurat_cluster.04, data = e6, FUN = mean )

Figure.3E3 = ggplot(values, aes(x = cell_cycle_all_strong1, y = Glycolytic1)) +
  geom_point(size = 3, shape = 16) +
  labs(y = "Glycolysis (HALLMARKS) Signature Score", x = "cell cycle Signature Score",
       title = "Average Signature Scores Per Seurat Cluster") +
theme_classic() +
    geom_text(aes(label = seurat_cluster.04), hjust = 0.5, vjust = -0.5, size = 8) +
    theme(axis.title.x = element_text(size = 14),
                axis.title.y = element_text(size = 14))
Figure.3E3
```

## Figure 3F
```{r}
fig4f.gene_list = c("HK2", "LDHA" ,"PFKM" ,"SLC2A1" ,"GAPDH",
"HIF1A" ,"IRF4", "RPIA", "TALDO1","CS", "IDH1", "IDH2", "SDHA", "FH",
"ATP5F1A", "CYCS", "CPT1A", "HADHA", "NRF1", "TFAM", "OPA1", "MFN1", "MFN2",
"DNM1L", "PPARG", "FABP4", "FABP5", "SLC3A2", "SLC7A5", "SLC1A5")
                                      
fig4f.aggregate_gene_list_df<- as.data.frame(AggregateExpression(no_tcr_final, features = fig4f.gene_list, group.by ="seurat_cluster.04", return.seurat = F, assays = "RNA"))

##organize the dataframe rows acording to gene_list
fig4f.aggregate_gene_list_df<- data.matrix(fig4f.aggregate_gene_list_df)[match(fig4f.gene_list , rownames(data.matrix(fig4f.aggregate_gene_list_df))),]



Idents(no_tcr_final) = "seurat_cluster.04"
colnames(fig4f.aggregate_gene_list_df) = c(levels(no_tcr_final))


scaled.mat4 <- t(scale(t(fig4f.aggregate_gene_list_df)))


Fig4f.mat = data.matrix(scaled.mat4)[match(fig4f.gene_list, rownames(data.matrix(scaled.mat4))),]
data.matrix(Fig4f.mat)

library(pheatmap)
input.hm.filter.breaks <- t(scale(t(Fig4f.mat )))
quantile.range <- quantile(t(input.hm.filter.breaks), probs = seq(0, 1, 0.1))
paletteBreaks <- quantile.range
paletteLength <- length(paletteBreaks)-1
col.pal <- RColorBrewer::brewer.pal(paletteLength, "RdBu")
col.pal <- rev(col.pal)

pheatmap(
    mat               = Fig4f.mat,
    breaks            = paletteBreaks,
    cluster_cols      = TRUE,
    cluster_rows      = FALSE,
    color             = col.pal,
    border_color      = NA,
    show_colnames     = T,
    show_rownames     = T,
    drop_levels       = TRUE,
    fontsize          = 14,
    main              = "fig4f",
     # Add this line to reorder the heatmap based on column
)
```